---
title: "Map of Houston Area Recycling Centers"
author: "Justin Z"
date: "June 4, 2019"
output: html_document
---


```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)

```


## Overview

This report is for the week 2 peer-graded project in the Developing Data
Products course from Johns Hopkins University within the Data Science
Specialization on Coursera. The instructions say to create a web page using R
Markdown that features a map created with Leaflet. 

To meet the objective a map was created which shows recycling locations in the
area of Houston, TX. The input for this document is a static file of the COHGIS
Recycling Center Locations dataset which came from the URL below. This is
essentially a reproduction of a feature that already exists on their website.
This report page is hosted by GitHub Pages, and the repo for this project is
located [here][2]

[2]: https://github.com/jtzingsheim1/Houston-Recycling-Map  "GitHub repo"

## Acknowledgement

Acknowledgement for this dataset goes to the City of Houston who make this data
available through their [GIS open data portal][1]

[1]: https://cohgis-mycity.opendata.arcgis.com/  "COHGIS Website"


## Part 1) Loading and Processing the Data

In the code chunk below the required packages are loaded and the raw data is
downloaded. The data are then processed to prepare for mapping.

```{r loading, message = FALSE}

# Load necessary packages
library(jsonlite)
library(tidyverse)
library(leaflet)

# Load in data through API
url1 <- "https://opendata.arcgis.com/datasets/2934190c5b8c4f508d80f22905d8cacc_2.geojson"
data.raw <- fromJSON(url1)$features  # 39 obs. of 3 variables
rec.data <- bind_cols(data.raw$properties, data.raw$geometry)  # 39 obs. of 9

# Split the coordinates column and add to data frame
coordinate.cols <- as.data.frame(do.call(rbind, rec.data$coordinates))  # 39 x 2
rec.data <- bind_cols(rec.data, coordinate.cols)  # 39 obs. of 11 variables
rm(url1, coordinate.cols)

# Remove unneeded columns and convert data types
rec.data <- rec.data %>%
  select(Longitude = V1, Latitude = V2, Location.Name = LOCATION,
         Location.Add = ADDRESS, Location.Status = Status,
         Location.Hours = Hours_1) %>%  # Drop and rename variables
  mutate(Location.Status = as.factor(Location.Status))  # 39 obs. of 6 variables

```


## Part 2) Generating the Map

In the next code chunk the data are placed into the map.

```{r mapping}

rec.data %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(lng = rec.data$Longitude, lat = rec.data$Latitude,
             popup = paste("<b>", rec.data$Location.Name, "</b><br>",
                           "Address: ", rec.data$Location.Add, "<br>",
                           "Status: ", rec.data$Location.Status, "<br>",
                           "Hours: ", rec.data$Location.Hours, sep = ""))

```

